---
title: "rollup"
author: "Ju Young Ahn"
date: "2024-08-27"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rollup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(rollup))
```

# rollup

`rollup`: A Tidy implementation of `GROUPING SETS`, `WITH ROLLUP`, and `WITH CUBE`, which are powerful extensions of the `GROUP BY` clause that compute multiple group-by clauses in a single statement in `SQL`. This package operates on top of the `dplyr` and performs the same functions as `SQL`.

## Usage
### Sample data
##### Description of sample data
- date_id : yyyy-mm-dd
- id : user unique id
- gender : male(M), female(F)
- age : age band (categorical)
- page_view_cnt : pageview count of user on date_id
- producte_view_cnt : product view count of user on date_id
```{r setup}
library(dplyr)
library(rollup)
data("web_service_data")
web_service_data %>% head
```

### grouping_sets
- `grouping_sets` is useful when multiple group-by operations are needed.
- `grouping_sets(a)` is equivalent to the single grouping set operation `group_by(a)`.
- `grouping_sets(a,b)` is equivalent to row binding of `group_by(a)` and `group_by(b)`.
- `grouping_sets((a,b),a,b, NA)` is equivalent to row binding of `group_by(a,b)`, `group_by(a)`, `group_by(b)` and without implementing the `group_by` operation. 
```{r warning=FALSE}
library(tidyr)
# compute average of `page_view_cnt` group by "gender", "age", and "gender & age", along with the overall average.
web_service_data %>% filter(date_id == '2024-06-30' & gender != "N") %>% 
  group_by(gender, age) %>% grouping_sets('gender', 'age', c('gender','age'), NA) %>% 
  summarize(avg_pv_cnt = mean(page_view_cnt))

# compute average of `page_view_cnt` group by "gender & age & product_view_cnt_cat" along with the marginal average with regard to "product_view_cnt_cat".
web_service_data %>% filter(date_id == '2024-06-30' & gender != "N") %>% 
  group_by(gender, age, product_view_cnt_cat) %>% grouping_sets('product_view_cnt_cat', c('product_view_cnt_cat', 'gender','age')) %>% 
  summarize(avg_pv_cnt = mean(page_view_cnt)) %>% pivot_wider(names_from = product_view_cnt_cat, values_from = avg_pv_cnt)
```

### with_cube
- `with_cube()` function is a simplified way of expressing `grouping_sets()`. 
- `with_cube()` is used when all possible combinations of variables are needed. 
- `group_by(a,b,c)` followed by `with_cube()` equals to `grouping_sets((a,b,c), (a,b), (a,c), (b,c), a, b, c, NA)`.
- `with_cube()` is particularly useful when you want to include total aggregates of both rows and columns in a cross table.

```{r warning=FALSE}
# compute and add total aggregates of rows and columns in 2x2 cross table using the `with_cube()`
web_service_data %>% filter(date_id == '2024-06-30' & gender != "N") %>% 
  group_by(gender, age) %>% with_cube() %>% 
  summarize(avg_pv_cnt = mean(page_view_cnt)) %>% pivot_wider(names_from = age, values_from = avg_pv_cnt)

# `with_cube()` equals to `grouping_sets()` with all possible combinations
# web_service_data %>% filter(date_id == '2024-06-30' & gender != "N") %>%
#   group_by(gender, age) %>% grouping_sets("gender","age",c("gender","age"), NA) %>%
#   summarize(avg_pv_cnt = mean(page_view_cnt)) %>% pivot_wider(names_from = age, values_from = avg_pv_cnt)
```

### with_rollup
- `with_rollup()` is useful when variables have a hierarchy, because all possible combinations are not necessary.
- `with_rollup()` equals to `grouping_sets()` with combinations in descending order.
- `group_by(a,b)` followed by `with_rollup()` equals to `grouping_sets((a,b), a, NA)`.
- `group_by(a,b,c)` followed by `with_rollup()` equals to `grouping_sets((a,b,c), (a,b), (a), NA)`.

```{r warning=FALSE}
# The variables "age_big" and "age" have a hierarchy. 
web_service_data_processed <- web_service_data %>% mutate(
  age_big = case_when(
    age %in% c(10,20,30) ~ 'young',
    age %in% c(40,50,60) ~ 'old'  
  )
)

# Therefore, if there are results aggregated by "age_big & age", marginal aggregates for "age" are not necessary.
# The following code calculates the aggregates for "age_big & age", marginal aggregates for "age_big", and aggregates for the entire data set.
web_service_data_processed %>% group_by(age_big, age) %>% with_rollup() %>% summarize(
  user_cnt = n_distinct(id),
  avg_pv_cnt = mean(page_view_cnt)
)

# `with_rollup()` equals to `grouping_sets()` with combinations in descending order.
# web_service_data_processed %>% group_by(age_big, age) %>% grouping_sets(c("age_big","age"),"age_big",NA) %>% summarize(
#   user_cnt = n_distinct(id),
#   avg_pv_cnt = mean(page_view_cnt)
# )
```
