---
title: "rollup"
author: "Ju Young Ahn"
date: "2024-07-30"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rollup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# rollup

`rollup`: A Tidy implementation of GROUPING SETS, WITH ROLLUP, and WITH CUBE, which are powerful extensions of the GROUP BY clause that compute multiple group-by clauses in a single statement in SQL. This package operates on top of the `dplyr` and performs the same functions as SQL.

## Usage
### Sample data
```{r setup}
library(dplyr)
library(rollup)
data("web_service_data")
web_service_data %>% head
```

### grouping_sets

- `grouping_sets(a)` is equivalent to the single grouping set operation `group_by(a)`.
- `grouping_sets(a,b)` is equivalent to row bind of `group_by(a)` and `group_by(b)`.
```{r warning=FALSE}
library(tidyr)
# avg_pv_cnt group by (gender, age, (gender, age))
web_service_data %>% filter(date_id == '2024-06-30' & gender != "N") %>% 
  group_by(gender, age) %>% grouping_sets('gender', 'age', c('gender','age')) %>% 
  summarize(avg_pv_cnt = mean(page_view_cnt))

# avg_pv_cnt group by ((gender, age, product_view_cnt_cat), product_view_cnt_cat)
web_service_data %>% filter(date_id == '2024-06-30' & gender != "N") %>% 
  group_by(gender, age, product_view_cnt_cat) %>% grouping_sets('product_view_cnt_cat', c('product_view_cnt_cat', 'gender','age')) %>% 
  summarize(avg_pv_cnt = mean(page_view_cnt)) %>% pivot_wider(names_from = product_view_cnt_cat, values_from = avg_pv_cnt)
```

### with_cube

- `with_cube()` adds all possible combinations of grouping variables
- `with_cube()` easily adds row sum and column sum in a cross table
- `group_by(a,b,c)` followed by `with_cube()` equals to `grouping_sets((a,b,c), (a,b), (a,c), (b,c), a, b, c, NA)`.

```{r warning=FALSE}
# add sub-total rows to 2x2 cross table using with_cube()
web_service_data %>% filter(date_id == '2024-06-30' & gender != "N") %>% 
  group_by(gender, age) %>% with_cube() %>% 
  summarize(avg_pv_cnt = mean(page_view_cnt)) %>% pivot_wider(names_from = age, values_from = avg_pv_cnt)

# with_cube equals to grouping_sets with all possible combinations
web_service_data %>% filter(date_id == '2024-06-30' & gender != "N") %>%
  group_by(gender, age) %>% grouping_sets("gender","age",c("gender","age"), NA) %>%
  summarize(avg_pv_cnt = mean(page_view_cnt)) %>% pivot_wider(names_from = age, values_from = avg_pv_cnt)
```

### with_rollup

- `with_rollup()` is an extension of the `group_by` clause that produces sub-total rows alongside grouped rows. 
- `group_by(a,b)` followed by `with_rollup()` equals to `grouping_sets((a,b), a, NA)`.
- `group_by(a,b,c)` followed by `with_rollup()` equals to `grouping_sets((a,b,c), (a,b), (a), NA)`.

```{r warning=FALSE}
web_service_data %>% 
  group_by(date_id) %>% with_rollup() %>% 
  summarize(user_cnt = n_distinct(if_else(page_view_cnt > 0, id, NA)))

# with_rollup equals to grouping_sets with all possible combinations in descending order
web_service_data %>%
  group_by(date_id) %>% grouping_sets("date_id", NA) %>%
  summarize(user_cnt = n_distinct(if_else(page_view_cnt > 0, id, NA)))
```