---
title: "rollup"
author: "Ju Young Ahn"
date: "2024-07-30"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rollup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# rollup

Grouping Sets, With Rollup and With Cube implementation for R dataframe

## Usage
### Sample data
```{r setup}
library(dplyr)
library(rollup)
data("web_service_data")
web_service_data %>% head
```

### grouping_sets

-   group by multiple columns in one line

```{r warning=FALSE}
# avg_pv_cnt group by (gender, age, (gender, age))
web_service_data %>% filter(date_id == '2024-06-30' & gender != "N") %>% 
  group_by(gender, age) %>% grouping_sets('gender', 'age', c('gender','age')) %>% 
  summarize(avg_pv_cnt = mean(page_view_cnt))

# avg_pv_cnt group by ((gender, age, product_view_cnt_cat), product_view_cnt_cat)
web_service_data %>% filter(date_id == '2024-06-30' & gender != "N") %>% 
  group_by(gender, age, product_view_cnt_cat) %>% grouping_sets('product_view_cnt_cat', c('product_view_cnt_cat', 'gender','age')) %>% 
  summarize(avg_pv_cnt = mean(page_view_cnt)) %>% pivot_wider(names_from = product_view_cnt_cat, values_from = avg_pv_cnt)
```

### with_cube

-   with_cube add all possible combinations of grouping variables
-   easily add row sum and column sum in cross table

```{r warning=FALSE}
web_service_data %>% filter(date_id == '2024-06-30' & gender != "N") %>% 
  group_by(gender, age) %>% with_cube() %>% 
  summarize(avg_pv_cnt = mean(page_view_cnt)) %>% pivot_wider(names_from = age, values_from = avg_pv_cnt)

# with_cube equals to grouping_sets with all possible combinations
# web_service_data %>% filter(date_id == '2024-06-30' & gender != "N") %>% 
#   group_by(gender, age) %>% grouping_sets("gender","age",c("gender","age"), NA) %>% 
#   summarize(avg_pv_cnt = mean(page_view_cnt)) %>% pivot_wider(names_from = age, values_from = avg_pv_cnt)
```

### with_rollup

-   easily add all possible combinations of grouping variables in descending order
-   calculate group sum and total sum (DAU and MAU in this example)

```{r warning=FALSE}
web_service_data %>% 
  group_by(date_id) %>% with_rollup() %>% 
  summarize(user_cnt = n_distinct(if_else(page_view_cnt > 0, id, NA)))

# with_rollup equals to grouping_sets with all possible combinations in descending order
# web_service_data %>%
#   group_by(date_id) %>% grouping_sets("date_id", NA) %>%
#   summarize(user_cnt = n_distinct(if_else(page_view_cnt > 0, id, NA)))
```

### sparklyr

-   All functions are also available in sparklyr (spark dataframe)
